Garbage Collection Evaluation Infrastructure for the Cog VM

### 2018-06-12 paper.tex
intro beginning

- Trying to build a more advanced full GC algo, need to measure..
- ref impl to compare
- Try to build benchmarks, both from real app and to stress specific aspects.
- We end by applying the bench on reference impl to validate the behavior is as expected.

@@

\section{Experimental results}
\label{sec:valid}

>>>>

The Cog virtual machine (VM), the default VM for multiple programming languages such Pharo, Squeak and Newspeak, currently features in production a stop-the-world Mark-Compact algorithm as the full garbage collector (GC) algorithm. The algorithm has a high throughput but a high pause time during which the application is not responsive (the GC interrupts the application for multiple seconds on modern Macbooks starting from multiple Gbs heaps). For interactive applications, a new algorithm is required with a smaller pause time.

Building and tuning a new GC algorithm requires to evaluate its behavior. A benchmarking infrastructure is required to do so. To build GC benchmarks, we took two approaches.

Firstly, we contacted multiple companies using the Cog VM in production on memory intensive application (>1Gb heaps) and built benchmarks out of their deployed application. As an example, we will discuss in the paper the Moose benchmarks implemented in collaboration with Feenk\footnote{https://feenk.com/}. Part of the business of Feenk consists in analysing software written in multiple programming languages using the open-source framework Moose~\cite{MoosePaper1,MooseBook1}. Using Moose, the application parses the software to analyse into a model, performs analysis on it and lastly release the model. This behavior was turned into three benchmarks, growing, accessing and shrinking the heap.

Secondly, \todo{Sophie :-) a few sentence on your bench}

- Gc algo ref

- We then evaluated the two reference algorithms on the infrastructure built to have reference benchmark results.

- related and futur

@@

\section{Reference Results}
\label{sec:valid}
